Analysis of Bauer et al. (unpublished) Field experiment: <br> Persistence

### 2022-12-07 R/model_persistence_check.Rmd
Update model_persistence_check.Rmd

```{r}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""),
  col_types = cols(
    .default = "?",
    plot = "f",
    site = "f",
    sand_ratio = "f",
    substrate_depth = col_factor(levels = c("30", "15")),
    target_type = col_factor(levels = c("hay_meadow", "dry_grassland")),
    seed_density = "f",
    exposition = col_factor(levels = c("north", "south")),
    survey_year = "c"
  )
  ) %>%
  ### Exclude data of seed mixtures
  filter(presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, exposition, sep = " "),
    botanist_year = factor(botanist_year),
    id = factor(id),
    n = persistence
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
  )
```

@@

```{r echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Sand ratio [vol%]")
plot2 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Substrate depth [cm]")
plot3 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Target type")
plot4 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Seed density [g/m²]")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Exposition")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Blocks")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

@@

```{r echo = FALSE}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

@@

```{r collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_persistence_2.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
# BARG 1.E
#load(file = here("outputs", "models", "model_persistence_2_prior.Rdata"))
# BARG 5.A/B/C
#load(file = here("outputs", "models", "model_persistence_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
```

@@

```{r}
get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site/plot) + (1 | botanist_year),
          data = sites)
ggplot(data = data.frame(x = c(0, 1)), aes(x = x)) +
  stat_function(fun = dbeta, n = 101, args = list(shape1 = 3.5, shape2 = 2.2)) +
  expand_limits(y = 0) +
  ggtitle("Beta distribution for Intercept")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = .2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution") # See Lemoine 2019 https://doi.org/10.1111/oik.05985
ggplot(data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student~italic(t)*"-distribution")) # Software standard
```

@@

```{r}
brms::prior_summary(m_1, all = FALSE)
```

@@

```{r collapse = TRUE}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

@@

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

@@

```{r collapse = TRUE}
mcmc_neff(neff_ratio(m_1))
mcmc_neff(neff_ratio(m_2))
```

@@

```{r}
mcmc_trace(posterior1, np = hmc_diagnostics1, facet_args = list(ncolumn = 2))
mcmc_trace(posterior2, np = hmc_diagnostics2, facet_args = list(ncolumn = 2))
mcmc_pairs(m_1, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
mcmc_pairs(m_2, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
mcmc_parcoord(posterior1, np = hmc_diagnostics1) +
  theme(axis.text.x = element_text(angle = 45))
mcmc_parcoord(posterior2, np = hmc_diagnostics2) +
  theme(axis.text.x = element_text(angle = 45))
```

@@

```{r}
p1 <- loo::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- loo::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2

@@

```{r max_height = "1000px"}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
```

```{r echo = FALSE}
write.csv(draws1, here("outputs", "statistics", "table_persistence.csv"))
```

@@

```{r echo = FALSE, max_height = "500px"}
sessionInfo()
```

>>>>

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""),
  col_types = cols(
    .default = "?",
    plot = "f",
    site = "f",
    sand_ratio = "f",
    substrate_depth = col_factor(levels = c("30", "15")),
    target_type = col_factor(levels = c("hay_meadow", "dry_grassland")),
    seed_density = "f",
    exposition = col_factor(levels = c("north", "south")),
    survey_year = "c"
  )
  ) %>%
  ### Exclude data of seed mixtures
  filter(presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, exposition, sep = " "),
    botanist_year = factor(botanist_year),
    id = factor(id),
    n = persistence
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
  )
```

@@

```{r data-exploration, echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Sand ratio [vol%]")
plot2 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Substrate depth [cm]")
plot3 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Target type")
plot4 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Seed density [g/m²]")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Exposition")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Blocks")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

@@

```{r outliers, echo = FALSE}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

@@

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_persistence_2.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
# BARG 1.E
#load(file = here("outputs", "models", "model_persistence_2_prior.Rdata"))
# BARG 5.A/B/C
#load(file = here("outputs", "models", "model_persistence_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
```

@@

```{r possible-priors}
get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site/plot) + (1 | botanist_year),
          data = sites)
ggplot(data = data.frame(x = c(0, 1)), aes(x = x)) +
  stat_function(fun = dbeta, n = 101, args = list(shape1 = 3.5, shape2 = 2.2)) +
  expand_limits(y = 0) +
  ggtitle("Beta distribution for Intercept")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = .2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution") # See Lemoine 2019 https://doi.org/10.1111/oik.05985
ggplot(data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student~italic(t)*"-distribution")) # Software standard
```

@@

```{r prior-summary}
brms::prior_summary(m_1, all = FALSE)
```

@@

```{r dharma, collapse = TRUE}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

@@

```{r analysis-preparation}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r rhat, collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

@@

```{r ess, collapse = TRUE}
mcmc_neff(neff_ratio(m_1))
mcmc_neff(neff_ratio(m_2))
```

@@

```{r mcmc}
mcmc_trace(posterior1, np = hmc_diagnostics1, facet_args = list(ncolumn = 2))
mcmc_trace(posterior2, np = hmc_diagnostics2, facet_args = list(ncolumn = 2))
mcmc_pairs(m_1, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
mcmc_pairs(m_2, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
mcmc_parcoord(posterior1, np = hmc_diagnostics1) +
  theme(axis.text.x = element_text(angle = 45))
mcmc_parcoord(posterior2, np = hmc_diagnostics2) +
  theme(axis.text.x = element_text(angle = 45))
```

@@

```{r loo-pit}
p1 <- loo::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- loo::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2

@@

```{r effect-sizes}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
```

```{r write-csv, echo = FALSE}
write.csv(draws1, here("outputs", "statistics", "table_persistence.csv"))
```

@@

```{r session-info, echo = FALSE}
sessionInfo()
```
